---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
# Analysis of daly activity
Physical activity is one of the most crucial aspects of a health lifestyle. Individuals engaged in enough exercise and activity are more likely to enjoy a life free of chronic conditions like diabetes.

## Loading and preprocessing the data
```{r, echo = TRUE, warning=FALSE}
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(anytime))
suppressPackageStartupMessages(library(lattice))

### Data Transformations

# create a temporary directory to store the downloaded file
temp <- tempfile()

# Downloading the file containing the required data
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", temp)

# Unzipping the downloaded file
con <- unz(temp, "activity.csv")

# Extracting the csv from the unzipped file
data <- read.table(con, header=T, sep = ",")
unlink(temp)
```


## What is mean total number of steps taken per day?
```{r, echo = TRUE, warning=FALSE}
# We need to omitt rows with na in "data"
df <- data %>% drop_na()
head(df)
#### What is mean total number of steps taken per day?
# 1. Total number of steps
Total <- df %>% 
    summarise(Tot_steps = sum(steps))
#mean_steps <- df %>% summarise(Tot_steps = round(sum(steps)/n(),1))
medmean <- df %>% 
  summarise(Aver = round(mean(steps)), Med = median(steps))
```

1. The total number of steps is `r Total`.
2. The mean number of steps is `r medmean$Aver` and the median number of steps is `r medmean$Med`.


## What is the average daily activity pattern?
```{r, echo = TRUE}
# 3. mean and median of total number of steps
average <- mean(Total$Tot_steps) 
med <- median(Total$Tot_steps)    

int_steps <- df %>% 
                group_by(interval) %>% 
                        summarise(average = mean(steps)) 

# A  Series plot of average number of steps by intervals 
plot(int_steps$interval, int_steps$average, type ="l",
     main = "Average Daily Activity Pattern", 
     xlab = "Intervals", ylab = "Average number of steps")

# Which 5-minute interval, on average across all the days in the dataset, 
#  contains the maximum number of steps?
max_steps <- int_steps %>% 
                        filter(average == max(average)) %>% 
                                       select(interval, average)
```

The daily activity pattern is almost normal with a median of `r round(med)` and a mean of `r round(average)`. The pattern has a maximum number of steps being `r round(max_steps$average)` at the interval `r max_steps$interval`

## Imputing missing values
```{r, echo = TRUE}
# Imputing missing values
## 1. Calculate and report the total number of missing values in 
##### the dataset (i.e. the total number of rows with NAs)
x_na <- sum(is.na(data$steps))

## 2-3. Filling in missing values with medians and creating a new dataset

df2 <- data 

# Devise a strategy for filling in all of the missing values in the datase.
# In place of NA, using the mean for that 5-minute interval
df2$steps[is.na(df2$steps)] <- with(df2, ave(steps, interval, 
                                          FUN = function(x) median(x, na.rm = TRUE)))[is.na(df2$steps)]
## Histogram
Tot_steps_miss <- df2 %>%
              group_by(date) %>%
                        summarise(Tot = sum(steps))
hist(Tot_steps_miss$Tot, main = "Daily Number of Steps",
     xlab="Steps", 
     border="red", 
     col="blue",
     las=1)
```

The data has `r x_na` rows with missing data.

## Are there differences in activity patterns between weekdays and weekends?
```{r, echo = TRUE}
## Weekdays and weekends 
df3 <- df2 %>% 
  mutate(day = weekdays(anytime(date)), Week = ifelse(day %in% c("Saturday","Sunday"), "Weekend", "Weekday"))
                 
df3 <- aggregate(df3$steps, by = list(df3$Week, df3$interval), mean, na.rm=TRUE)
names(df3) <- c("Week", "interval", "steps") 

average1 <- mean(df3$steps) 
med1 <- median(df3$steps) 
 

## Creating a panel plot
p <- xyplot(steps ~ interval | factor(Week), data=df3, 
            type = 'l',
            main="Average Number of Steps by type of weekday",
            xlab="Time Intervals",
            ylab="Average Number of Steps", cex.main=2.5)
print (p)  
```

The mean number of steps is `r round(average1)` and the median number of steps is `r round(med1)`. Imputation of missing values has an effect on the mean and median. When missing data was removed, the mean was 37 and the median 0.